---
title: "General statistical overviews of FILTER data"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    code_folding: hide
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,dpi=72,fig.retina=2,fig.width=8)
source(here::here("code/common_basis.R"), local = knitr::knit_global())
```

# Temporal overview

```{r temporal_overview, fig.width=8, fig.height=8}
p_year %>% 
  inner_join(poems,by=c("p_id")) %>%
  count(collection,year) %>%
  mutate(measure="yearly count") %>%
  union_all(
    p_year %>% # 10 year rolling mean
    distinct(year) %>% 
    left_join(p_year %>% distinct(year),sql_on="RHS.year BETWEEN LHS.year-5 AND LHS.year+5") %>%
    inner_join(p_year,by=c("year.y"="year")) %>%
    inner_join(poems,by=c("p_id")) %>%
    group_by(collection=collection,year=year.x) %>%
    summarize(n=n()/10,.groups="drop") %>%
    mutate(measure="10 year rolling mean")
  ) %>%
  filter(year>0,year<9999,collection!="literary") %>%
  ggplot(aes(x=year,y=n,color=measure)) +
  geom_point(data=~.x %>% filter(measure=="yearly count")) +
  geom_line(data=~.x %>% filter(measure=="10 year rolling mean")) +
  theme_hsci_discrete(base_family="Arial") + 
  theme(legend.justification=c(0,1), legend.position=c(0.02, 0.98), legend.background = element_blank(), legend.key=element_blank()) + 
  labs(color=NULL) +
  scale_y_continuous(breaks=seq(0,20000,by=2000),labels=scales::comma_format()) +
  ylab("Poems") +
  scale_x_continuous(breaks=seq(1000,2000,by=50)) +
  xlab("Year") +
  facet_wrap(~collection, ncol=1) +
  ggtitle("Number of poems by year and collection")
```

```{r}
p_year %>% 
  filter(year %in% c(0,9999)) %>% 
  left_join(poems) %>% 
  count(collection,year) %>%
  ungroup() %>%
  gt() %>%
  tab_header(title="Abnormal years") %>%
  fmt_integer(n)
```

# Overview of collectors

```{r collectors_overview, fig.width=8, fig.height=11}
poems %>% 
  distinct(collection) %>%
  pull() %>%
  map(~p_col %>% 
    inner_join(poems %>% filter(collection==.x),by=c("p_id")) %>%
    count(col_id) %>%
    left_join(collectors,by=c("col_id")) %>%
    select(col_id,name,n) %>%
    collect() %>%
    mutate(col_id=fct_reorder(str_c(col_id,"|",name),n)) %>%
    mutate(col_id=fct_lump_n(col_id,n=100,w=n)) %>%
    mutate(col_id=fct_relevel(col_id,"Other")) %>%
    group_by(col_id) %>%
    tally(wt=n) %>% {
      ggplot(.,aes(x=col_id,y=n)) +
      geom_col() +
      geom_text(aes(label=p(n)),hjust='left',nudge_y = 100) +
      theme_hsci_discrete(base_family="Arial") +
      coord_flip() +
      labs(title=str_c("Collectors in ",.x))
    }
  )
```

```{r}
p_col %>% 
  anti_join(collectors) %>%
  count(col_id) %>%
  gt() %>%
  tab_header(title="Collectors without a name") %>%
  fmt_integer(n)
```

# Geographical overview

```{r}
d <- p_loc %>% 
  count(loc_id) %>% 
  inner_join(locations) %>%
  select(name,n) %>%
  collect()

poems_without_location <- poems %>% 
  anti_join(p_loc) %>% 
  count() %>% 
  pull()

unprojected_locations <- d %>%
  anti_join(polygons) %>%
  add_row(name=NA,n=poems_without_location)
```


```{r, fig.height=11}
polygons %>%
  left_join(d) %>%
  tm_shape() +
  tm_polygons(col='n', id='name', style='fisher', palette='plasma') +
  tm_layout(title=str_c("Geographical overview. Missing ",unprojected_locations %>% tally(wt=n) %>% pull() %>% p," poems."))
```

## Poem locations not mapped

```{r}
unprojected_locations %>%
  arrange(desc(n)) %>%
  gt() %>%
  tab_header("Poem locations not mapped") %>%
  fmt_integer(n)
```

## Geographical overview by collection

```{r}
d <- p_loc %>% 
  left_join(poems) %>%
  count(collection,loc_id) %>% 
  ungroup() %>%
  inner_join(locations) %>%
  select(collection,name,n) %>%
  collect()

poems_without_location <- poems %>% 
  anti_join(p_loc) %>% 
  count(collection) %>% 
  collect() %>%
  mutate(name=NA_character_)

unprojected_locations <- d %>%
  anti_join(polygons) %>%
  union_all(poems_without_location)
```

```{r, fig.height=11}
poems %>% 
  distinct(collection) %>%
  pull() %>%
  map(~
    tm_shape(
      polygons %>%
        left_join(
          p_loc %>% 
            inner_join(poems %>% filter(collection==.x),by=c("p_id")) %>%
            count(loc_id) %>% 
            inner_join(locations) %>%
            select(name,n) %>%
            collect()
        )
    ) +
    tm_polygons(col='n', id='name', style='fisher', palette='plasma') +
    tm_layout(title=str_c("Geography of ",.x,". Missing ",unprojected_locations %>% filter(collection==.x) %>% tally(wt=n) %>% pull() %>% p," poems."))
  )
```

## Poem locations not mapped by collection

```{r, results="asis"}
poems %>% 
  distinct(collection) %>%
  pull() %>%
  map(~
    unprojected_locations %>%
      filter(collection==.x) %>%
      arrange(desc(n)) %>%
      select(-collection) %>%
      gt() %>%
      tab_header(str_c("Poem locations not mapped in ",.x)) %>%
      fmt_integer(n)
  )
```

# Spatiotemporal overview

```{r}
d <- poems %>%
  left_join(p_year %>% mutate(year=if_else(year %in% c(0L,9999L),NA,year))) %>% 
  collect() %>%
  mutate(year_ntile=ntile(year,11)) %>%
  group_by(year_ntile) %>%
  mutate(years=str_c(min(year),"-",max(year))) %>%
  ungroup() %>%
  left_join(p_loc %>% collect()) %>% 
  count(years,loc_id) %>% 
  ungroup() %>%
  left_join(locations %>% select(loc_id,name) %>% collect())
```


```{r,fig.height=11}
polygons %>% 
  left_join(d %>% complete(name,years)) %>%
  tm_shape() +
  tm_polygons(col='n', id='name', style='fisher', palette='plasma') +
  tm_layout(main.title="Geographical overviews by time",legend.outside.size=0.1) +
  tm_facets(by="years",ncol=4)
```

# Poem verse statistics

## Line types

```{r}
d <- verses %>% 
  left_join(verse_poem) %>% 
  left_join(poems) %>% 
  count(collection,type) %>% 
  ungroup() %>%
  arrange(collection,desc(n)) %>%
  collect()
```


```{r}
d %>% 
  group_by(collection) %>%
  mutate(proportion=n/sum(n)) %>%
  gt() %>%
  fmt_integer(n) %>%
  fmt_percent(proportion)
```

## Verse line lengths

```{r}
d_nr_characters <- verses_cl %>%
  mutate(nr_characters=str_length(text)) %>%
  left_join(verse_poem) %>% 
  left_join(poems) %>% 
  count(collection,nr_characters) %>% 
  ungroup() %>%
  arrange(collection,desc(n)) %>%
  collect()

d_nr_words <- word_occ %>%
  group_by(v_id) %>%
  summarise(nr_words=max(pos),.groups="drop") %>%
  left_join(verse_poem) %>%
  left_join(poems) %>% 
  count(collection,nr_words) %>% 
  ungroup() %>%
  arrange(collection,desc(n)) %>%
  collect()
```
### Verse line lengths in characters
```{r}
d_nr_characters %>% 
  filter(nr_characters<=60) %>%
  ggplot(aes(x=nr_characters,y=n)) +
  geom_col(width=1) +
  facet_wrap(~collection,scales="free_y") +
  theme_hsci_discrete(base_family="Arial") +
  scale_y_continuous(labels=scales::comma_format()) +
  xlab("Number of characters") +
  ylab("Verses") +
  labs(title="Number of characters in verse lines")
```

```{r}
d_nr_characters %>% 
  group_by(collection) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nr_characters<=60) %>%
  ggplot(aes(x=nr_characters,y=collection,fill=collection,height=prop)) +
  geom_density_ridges(stat='identity') +
  theme_hsci_discrete(base_family="Arial") +
#  scale_y_continuous(labels=scales::percent_format()) +
  xlab("Number of characters") +
  ylab("Verses") +
  labs(title="Number of characters in verse lines")
```


### Verse lines with more than 60 characters

```{r}
d_nr_characters %>% 
  mutate(nl=if_else(nr_characters>60,n,0L)) %>%
  group_by(collection) %>%
  summarise(lines=sum(nl),proportion=sum(nl)/sum(n),.groups="drop") %>%
  arrange(desc(lines)) %>%
  gt() %>%
  tab_header(title="Verse lines with more than 60 characters") %>%
  fmt_integer(lines) %>%
  fmt_percent(proportion)
```


### Verse line lengths in words
```{r}
d_nr_words %>% 
  filter(nr_words<=10) %>%
  ggplot(aes(x=nr_words,y=n)) +
  geom_col(width=1) +
  facet_wrap(~collection,scales="free_y") +
  scale_x_continuous(breaks=seq(0,10,by=2)) +
  scale_y_continuous(labels=scales::comma_format()) +
  theme_hsci_discrete(base_family="Arial") +
  xlab("Number of words") +
  ylab("Verses") +
  labs(title="Number of words in verse lines")
```

```{r}
d_nr_words %>% 
  filter(nr_words<=10) %>%
  uncount(n) %>%
  ggplot(aes(x=nr_words,y=collection,fill=collection)) +
  stat_binline(binwidth=1) +
  theme_hsci_discrete(base_family="Arial") +
  scale_x_continuous(breaks=seq(0,10,by=2)) +
  xlab("Number of words") +
  ylab("Verses") +
#  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Number of words in verse lines")
```

### Verse lines with more than 10 words

```{r}
d_nr_words %>% 
  mutate(nl=if_else(nr_words>10,n,0L)) %>%
  group_by(collection) %>%
  summarise(lines=sum(nl),proportion=sum(nl)/sum(n),.groups="drop") %>%
  arrange(desc(lines)) %>%
  gt() %>%
  tab_header(title="Verse lines with more than 10 words") %>%
  fmt_integer(lines) %>%
  fmt_percent(proportion)
```

```{r}
verse_nr_words <- word_occ %>% 
  group_by(v_id) %>%
  summarise(nr_words=max(pos)) %>%
  compute_a(unique_indexes=list(c("v_id","nr_words")))

word_nr_characters <- words %>%
  mutate(nr_characters=str_length(text)) %>%
  select(w_id,nr_characters) %>%
  compute_a(unique_indexes=list(c("w_id","nr_characters")))

d <- word_occ %>%
  left_join(word_nr_characters) %>%
  left_join(verse_nr_words) %>%
  left_join(verse_poem %>% select(-pos),by=c("v_id")) %>% 
  left_join(poems) %>% 
  count(collection,nr_words,pos,nr_characters) %>%
  collect()
```

```{r}
d %>%
  group_by(collection,nr_words,pos) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nr_words>=2,nr_words<6) %>%
  mutate(nr_words=as_factor(nr_words),pos=as_factor(pos)) %>%
  uncount(n) %>%
  ggplot(aes(x=nr_characters,y=nr_words,fill=nr_words)) +
  stat_binline(binwidth=1) +
  facet_grid(collection~pos,labeller = labeller(pos=label_both)) + 
  xlab("Number of characters in word") +
  ylab("Number of words in verse") +
  labs(
    title="Number of characters in words by their position",
    subtitle="According to length of verse and collection"
    ) +
  theme_hsci_discrete(base_family="Arial")
```

