---
title: "General statistical overviews of FILTER data"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    code_folding: hide
    toc: yes
  html_document:
    code_folding: hide
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,dpi=300,fig.retina=2,fig.width=8)
source(here::here("src/common_basis.R"))
tmap_mode("plot")
```

# Temporal overview

```{r}
p_year %>% 
  inner_join(poems,by=c("p_id")) %>%
  count(collection,year) %>%
  mutate(measure="yearly count") %>%
  union_all(
    p_year %>% # 10 year rolling mean
      distinct(year) %>% 
      left_join(p_year %>% distinct(year),sql_on="RHS.year BETWEEN LHS.year-5 AND LHS.year+4") %>%
      inner_join(p_year,by=c("year.y"="year")) %>%
      inner_join(poems,by=c("p_id")) %>%
      group_by(collection=collection,year=year.x) %>%
      summarize(n=n()/n_distinct(year.y),.groups="drop") %>%
      mutate(measure="10 year rolling mean")
  ) %>%
  filter(collection!="literary",!year %in% c(0,9999)) %>%
  mutate(year=if_else(year>=1800,year,1780)) %>%
  group_by(collection,measure,year) %>%
  summarise(n=sum(n),.groups="drop") %>%
  collect() %>%
  complete(year,collection,measure,fill=list(n=0)) %>%
  mutate(collection=fct_relevel(str_to_upper(collection),"ERAB","SKVR","JR")) %>%
  group_by(collection,measure) %>%
  arrange(year) %>%
  filter(n!=0 | lag(n)!=0 | lead(n)!=0) %>%
  ungroup() %>%
  mutate(youtlier=n>4600,xoutlier=year<1800) %>%
  ggplot(aes(x=year,y=n,color=collection)) +
  geom_point(data=~.x %>% filter(measure=="yearly count",youtlier==FALSE),size=0.5) +
  geom_point(data=~.x %>% filter(youtlier==TRUE),aes(x=year),y=5000) +
  geom_text_repel(data=~.x %>% filter(youtlier==TRUE),aes(x=year,label=scales::number(n)),y=5000, show.legend=FALSE) +
  geom_point(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=year,y=n)) +
  geom_text_repel(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=year,y=n,label=scales::number(n)), show.legend=FALSE) +  
  geom_line(data=~.x %>% filter(xoutlier==FALSE,measure=="10 year rolling mean")) +
  theme_hsci_discrete(base_family="Arial") + 
  theme(
    legend.justification=c(0,1), 
    legend.position=c(0.02, 0.98), 
    legend.background = element_blank(), 
    legend.key=element_blank()
  ) + 
  labs(color=NULL) +
  coord_cartesian(ylim=c(0,4600),xlim=c(1800,1970),clip="off") +
  scale_y_continuous(breaks=seq(0,20000,by=1000),labels=scales::number) +
#  ylab("Poems") +
  ylab("Runojen määrä") +
  scale_x_continuous(breaks=seq(1000,2000,by=10)) +
#  xlab("Year") +
  xlab("Vuosi") +
  ggtitle("")
#  ggtitle("Runojen määrä vuosittain ja kokoelmittain")
#  ggtitle("Number of poems by year and collection")
```

```{r}
top_top_themes <- poem_theme %>% 
  inner_join(poems) %>%
  inner_join(themes_to_top_level_themes) %>%
  count(collection, ancestor_t_id) %>%
  group_by(collection) %>%
  slice_max(n,n=9) %>%
  ungroup() %>% 
  mutate(top_theme=TRUE) %>%
  select(ancestor_t_id,top_theme) %>%
  compute_a(temporary=TRUE, overwrite=TRUE)
```


```{r}
d <- p_year %>% 
  inner_join(poems,by=c("p_id")) %>%
  inner_join(poem_theme %>%
    inner_join(themes_to_top_level_themes %>% 
                 inner_join(themes %>%
                              filter(!str_detect(theme_id,"^erab_orig")) %>%
                        select(ancestor_t_id=t_id,ancestor_theme_name=name)))) %>%
  left_join(top_top_themes) %>%
  mutate(
    ancestor_theme_name=if_else(!is.na(top_theme),ancestor_theme_name,"Muut"),
    ancestor_t_id=if_else(!is.na(top_theme),ancestor_t_id,-1),
    ) %>%
  replace_na(list(ancestor_theme_name="Tuntematon", ancestor_t_id=-2)) %>%
  distinct(ancestor_t_id,ancestor_theme_name, collection, year, p_id) %>%
  count(ancestor_t_id,ancestor_theme_name, collection, year) %>%
  mutate(measure="yearly count") %>%
  union_all(
    p_year %>% # 10 year rolling mean
      distinct(year) %>% 
      left_join(p_year %>% distinct(year),sql_on="RHS.year BETWEEN LHS.year-5 AND LHS.year+4") %>%
      inner_join(p_year,by=c("year.y"="year")) %>%
      inner_join(poems,by=c("p_id")) %>%
      inner_join(poem_theme %>%
        inner_join(themes_to_top_level_themes %>% 
                     inner_join(themes %>% 
                                  filter(!str_detect(theme_id,"^erab_orig")) %>% 
                                  select(ancestor_t_id=t_id,ancestor_theme_name=name)))) %>%
      left_join(top_top_themes) %>%
      mutate(
        ancestor_theme_name=if_else(!is.na(top_theme),ancestor_theme_name,"Muut"),
        ancestor_t_id=if_else(!is.na(top_theme),ancestor_t_id,-1),
        ) %>%
      replace_na(list(ancestor_theme_name="Tuntematon", ancestor_t_id=-2)) %>%
      distinct(ancestor_t_id,ancestor_theme_name, collection, year.x, year.y, p_id) %>%
      group_by(ancestor_t_id,ancestor_theme_name, collection, year=year.x) %>%
      summarize(n=n()/n_distinct(year.y),.groups="drop") %>%
      mutate(measure="10 year rolling mean")
  ) %>%
  filter(collection!="literary",!year %in% c(0L,9999L)) %>%
  mutate(year=if_else(year>=1800L,year,1780L)) %>%
  group_by(ancestor_theme_name, collection, measure, year) %>%
  summarise(n=sum(n),.groups="drop") %>%
  collect()
```

```{r}
d %>%
  mutate(collection=fct_relevel(str_to_upper(collection),"SKVR","ERAB","JR")) %>%
  filter(collection=="SKVR") %>%
  complete(ancestor_theme_name, year,collection,measure,fill=list(n=0)) %>%
  group_by(ancestor_theme_name, collection,measure) %>%
  arrange(year) %>%
  filter(n!=0 | lag(n)!=0 | lead(n)!=0) %>%
  ungroup() %>%
  mutate(youtlier=n>1300,xoutlier=year<1800) %>%
  ggplot(aes(x=year,y=n,color=ancestor_theme_name)) +
#  facet_wrap(~collection) +
  geom_point(data=~.x %>% filter(measure=="yearly count",youtlier==FALSE,xoutlier==FALSE),size=0.5) +
  geom_point(data=~.x %>% filter(youtlier==TRUE),aes(x=year),y=1400) +
  geom_text_repel(data=~.x %>% filter(youtlier==TRUE),aes(x=year,label=scales::number(n)),y=1400, show.legend=FALSE) +
  geom_point(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=1785,y=n)) +
  geom_text_repel(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=1785,y=n,label=scales::number(n)), show.legend=FALSE) +  
  geom_line(data=~.x %>% filter(xoutlier==FALSE,measure=="10 year rolling mean")) +
  theme_hsci_discrete(base_family="Arial") + 
  theme(
    legend.justification=c(0,1), 
    legend.position=c(0.02, 0.98), 
    legend.background = element_blank(), 
    legend.key=element_blank()
  ) + 
  labs(color=NULL) +
  coord_cartesian(ylim=c(0,1300),xlim=c(1800,1940),clip="off") +
  scale_y_continuous(breaks=seq(0,20000,by=500),labels=scales::number) +
#  ylab("Poems") +
  ylab("Runojen määrä") +
  scale_x_continuous(breaks=seq(1000,2000,by=10)) +
#  xlab("Year") +
  xlab("Vuosi") +
  ggtitle("")
#  ggtitle("Runojen määrä vuosittain ja kokoelmittain")
#  ggtitle("Number of poems by year and collection")
```

```{r}
d %>%
  mutate(collection=fct_relevel(str_to_upper(collection),"SKVR","ERAB")) %>%
  filter(collection=="ERAB") %>%
  complete(ancestor_theme_name, year,collection,measure,fill=list(n=0)) %>%
  group_by(ancestor_theme_name, collection,measure) %>%
  arrange(year) %>%
  filter(n!=0 | lag(n)!=0 | lead(n)!=0) %>%
  ungroup() %>%
  mutate(youtlier=n>820,xoutlier=year<1800) %>%
  mutate(ancestor_theme_name=case_match(ancestor_theme_name,
    "Laulud noorrahva elust" ~ "Laulut nuorison elämästä (Laulud noorrahva elust)",
    "Muut" ~ "Muut (sisältää 17 luokkaa)",
    "Laulud meelelahutamiseks" ~ "Viihdytyslaulut (Laulud meelelahutamiseks)",
    "Lüroeepilised laulud" ~ "Lyroeeppiset laulut (Lüroeepilised laulud)",
    "Laulud laulust" ~ "Laulut laulusta (Laulud laulust)",
    "Töölaulud" ~ "Työlaulut (Töölaulud)",
    "Looduslaulud" ~ "Laulut luonnosta (Looduslaulud)",
    "Laulud ühiskondlikest vahekordadest" ~ "Laulut yhteiskunnallisista suhteista\n(Laulud ühiskondlikest vahekordadest)",
    "Murelaulud" ~ "Huolilaulut (Murelaulud)",
    "Laulud abielust" ~ "Laulut avioelämästä (Laulud abielust)",
    "Kalendrilaulud" ~ "Kalendaarilaulut (Kalendrilaulud)"
  )) %>%
  mutate(ancestor_theme_name=fct_reorder(ancestor_theme_name,n,.fun=sum,.desc=TRUE)) %>%
  mutate(ancestor_theme_name=fct_relevel(ancestor_theme_name, "Muut (sisältää 17 luokkaa)", after=Inf)) %>%
  ggplot(aes(x=year,y=n,color=ancestor_theme_name)) +
#  facet_wrap(~collection) +
  geom_point(data=~.x %>% filter(measure=="yearly count",youtlier==FALSE,xoutlier==FALSE),size=0.5) +
  geom_point(data=~.x %>% filter(youtlier==TRUE),aes(x=year),y=900) +
  geom_text_repel(data=~.x %>% filter(youtlier==TRUE),aes(x=year,label=scales::number(n)),y=900, show.legend=FALSE) +
  geom_point(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=1785,y=n)) +
  geom_text_repel(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=1785,y=n,label=scales::number(n)), show.legend=FALSE) +  
  geom_line(data=~.x %>% filter(xoutlier==FALSE,measure=="10 year rolling mean")) +
  theme_hsci_discrete(base_family="Arial") + 
  theme(
    legend.justification=c(0,1), 
    legend.position=c(0.02, 0.98), 
    legend.background = element_blank(), 
    legend.key=element_blank()
  ) + 
  labs(color=NULL) +
  coord_cartesian(ylim=c(0,820),xlim=c(1820,1950),clip="off") +
  scale_y_continuous(breaks=seq(0,20000,by=200),labels=scales::number) +
#  ylab("Poems") +
  ylab("Runojen määrä") +
#  guides(color=guide_legend(nrow=2)) +
  scale_x_continuous(breaks=seq(1000,2000,by=10)) +
#  xlab("Year") +
  xlab("Vuosi") +
  ggtitle("")
#  ggtitle("Runojen määrä vuosittain ja kokoelmittain")
#  ggtitle("Number of poems by year and collection")
```

```{r}
d %>%
  mutate(collection=fct_relevel(str_to_upper(collection),"SKVR","ERAB","JR")) %>%
  filter(collection=="JR") %>%
  complete(ancestor_theme_name, year,collection,measure,fill=list(n=0)) %>%
  group_by(ancestor_theme_name, collection,measure) %>%
  arrange(year) %>%
  filter(n!=0 | lag(n)!=0 | lead(n)!=0) %>%
  ungroup() %>%
  mutate(youtlier=n>6500,xoutlier=year<1800) %>%
  ggplot(aes(x=year,y=n,color=ancestor_theme_name)) +
#  facet_wrap(~collection) +
  geom_point(data=~.x %>% filter(measure=="yearly count",youtlier==FALSE),size=0.5) +
  geom_point(data=~.x %>% filter(youtlier==TRUE),aes(x=year),y=7200) +
  geom_text_repel(data=~.x %>% filter(youtlier==TRUE),aes(x=year,label=scales::number(n)),y=7200, show.legend=FALSE) +
  geom_point(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=year,y=n)) +
  geom_text_repel(data=~.x %>% filter(xoutlier==TRUE,measure=="yearly count"),aes(x=year,y=n,label=scales::number(n)), show.legend=FALSE) +  
  geom_line(data=~.x %>% filter(xoutlier==FALSE,measure=="10 year rolling mean")) +
  theme_hsci_discrete(base_family="Arial") + 
  theme(
    legend.justification=c(0,1), 
    legend.position=c(0.02, 0.98), 
    legend.background = element_blank(), 
    legend.key=element_blank()
  ) + 
  labs(color=NULL) +
  coord_cartesian(ylim=c(0,6500),xlim=c(1800,1960),clip="off") +
  scale_y_continuous(breaks=seq(0,20000,by=1000),labels=scales::number) +
#  ylab("Poems") +
  ylab("Runojen määrä") +
  scale_x_continuous(breaks=seq(1000,2000,by=10)) +
#  xlab("Year") +
  xlab("Vuosi") +
  ggtitle("")
#  ggtitle("Runojen määrä vuosittain ja kokoelmittain")
#  ggtitle("Number of poems by year and collection")
```


```{r}
p_year %>% 
  filter(year %in% c(0,9999)) %>% 
  left_join(poems) %>% 
  count(collection,year) %>%
  ungroup() %>%
  gt() %>%
  tab_header(title="Abnormal years") %>%
  fmt_integer(n)
```

# Overview of collectors

```{r collectors_overview, fig.width=8, fig.height=11}
poems %>% 
  distinct(collection) %>%
  pull() %>%
  map(~p_col %>% 
    inner_join(poems %>% filter(collection==.x),by=c("p_id")) %>%
    count(col_id) %>%
    left_join(collectors,by=c("col_id")) %>%
    select(col_id,name,n) %>%
    collect() %>%
    mutate(col_id=fct_reorder(str_c(col_id,"|",name),n)) %>%
    mutate(col_id=fct_lump_n(col_id,n=100,w=n)) %>%
    mutate(col_id=fct_relevel(col_id,"Other")) %>%
    group_by(col_id) %>%
    tally(wt=n) %>% {
      ggplot(.,aes(x=col_id,y=n)) +
      geom_col() +
      geom_text(aes(label=p(n)),hjust='left',nudge_y = 100) +
      theme_hsci_discrete(base_family="Arial") +
      coord_flip() +
      labs(title=str_c("Collectors in ",.x))
    }
  )
```

```{r}
p_col %>% 
  anti_join(collectors) %>%
  count(col_id) %>%
  gt() %>%
  tab_header(title="Collectors without a name") %>%
  fmt_integer(n)
```

```{r}
p_col %>%
  inner_join(collectors) %>%
  inner_join(poems) %>%
  filter(collection!="literary") %>%
  mutate(collection=str_to_upper(collection)) %>%
  count(collection,col_id,name) %>%
  group_by(collection) %>%
  slice_max(order_by=n,n=10) %>%
  ungroup() %>%
  select(-col_id) %>%
  gt(groupname_col = "collection", rowname_col = "name") %>%
  row_group_order(c("ERAB","SKVR","JR")) %>%
  fmt_integer(n,sep_mark=" ")
```

```{r,fig.height=3}
p_col %>%
  inner_join(collectors) %>%
  inner_join(poems) %>%
  filter(collection!="literary") %>%
  mutate(collection=str_to_upper(collection)) %>%
  count(collection,col_id,name) %>%
  group_by(collection) %>%
  slice_max(order_by=n,n=10) %>%
  ungroup() %>%
  inner_join(p_col) %>%
  inner_join(poem_theme) %>%
  inner_join(themes_to_top_level_themes) %>%
  count(collection, col_id, t_id=ancestor_t_id) %>%
  inner_join(themes %>% rename(theme_name=name)) %>%
  inner_join(collectors) %>%
  filter(collection=="SKVR") %>%
  collect() %>% 
#  group_by(collection) %>%
#  mutate(theme_name=fct_lump_n(theme_name, n, n=5, other_level="Muut")) %>%
#  ungroup() %>%
  group_by(col_id) %>%
  mutate(tn=sum(n)) %>%
  ungroup() %>%
  mutate(name=fct_reorder(name,tn)) %>%
  ggplot(aes(x=name,fill=theme_name,y=n)) +
#  facet_wrap(~collection,scales="free",ncol=1) +
  geom_col() +
  theme_hsci_discrete() +
  coord_flip() +
  scale_y_continuous(labels=scales::number) +
  labs(fill="Päätyyppi") +
  xlab("Kerääjä") +
  ylab("Tyyppimerkintöjä")

```


# Geographical overview

```{r}
d <- p_loc %>% 
  count(loc_id) %>% 
  inner_join(locations) %>%
  select(name,n) %>%
  collect()

poems_without_location <- poems %>% 
  anti_join(p_loc) %>% 
  count() %>% 
  pull()

unprojected_locations <- d %>%
  anti_join(polygons) %>%
  add_row(name=NA,n=poems_without_location)
```


```{r, fig.height=11,results="asis"}
polygons %>%
  left_join(d) %>%
  tm_shape() +
  tm_polygons(col='n', id='name', style='fisher', palette='plasma') +
  tm_layout(title=str_c("Geographical overview. Missing ",unprojected_locations %>% tally(wt=n) %>% pull() %>% p," poems."))
```

## Poem locations not mapped

```{r}
unprojected_locations %>%
  arrange(desc(n)) %>%
  gt() %>%
  tab_header("Poem locations not mapped") %>%
  fmt_integer(n)
```

## Geographical overview by collection

```{r}
d <- p_loc %>% 
  left_join(poems) %>%
  count(collection,loc_id) %>% 
  ungroup() %>%
  inner_join(locations) %>%
  select(collection,name,n) %>%
  collect()

poems_without_location <- poems %>% 
  anti_join(p_loc) %>% 
  count(collection) %>% 
  collect() %>%
  mutate(name=NA_character_)

unprojected_locations <- d %>%
  anti_join(polygons) %>%
  union_all(poems_without_location)
```

```{r, fig.height=11}
poems %>% 
  distinct(collection) %>%
  pull() %>%
  map(~
    tm_shape(
      polygons %>%
        left_join(
          p_loc %>% 
            inner_join(poems %>% filter(collection==.x),by=c("p_id")) %>%
            count(loc_id) %>% 
            inner_join(locations) %>%
            select(name,n) %>%
            collect()
        )
    ) +
    tm_polygons(col='n', id='name', style='fisher', palette='plasma') +
    tm_layout(title=str_c("Geography of ",.x,". Missing ",unprojected_locations %>% filter(collection==.x) %>% tally(wt=n) %>% pull() %>% p," poems."))
  )
```

## Poem locations not mapped by collection

```{r, results="asis"}
poems %>% 
  distinct(collection) %>%
  pull() %>%
  map(~
    unprojected_locations %>%
      filter(collection==.x) %>%
      arrange(desc(n)) %>%
      select(-collection) %>%
      gt() %>%
      tab_header(str_c("Poem locations not mapped in ",.x)) %>%
      fmt_integer(n)
  )
```

# Informants

```{r}
raw_meta %>% 
  filter(field=="INF") %>%
  mutate(value_c=str_replace_all(value,"^\\s*[A-Za-zÅÄÖåäö][a-zåäö][a-zåäö]+\\.","")) %>% 
  mutate(name=str_replace_all(value_c,"\\s*\"*([A-Za-zÅÄÖåäö]?[a-zåäö]?[a-zåäö]?\\.?[^;.,]+)(.|\\n)*","\\1")) %>% 
  group_by(name) %>% 
  summarise(origs=str_flatten(sql("distinct value"),collapse="|"),n=n(),.groups="drop") %>% 
  collect()
```


# Poem types

```{r}
poems %>% 
  filter(collection!="literary") %>%
  left_join(poem_theme %>% filter(is_minor==0) %>% inner_join(themes %>% mutate(theme_type=if_else(str_detect(theme_id,"^erab_orig"),"Non-unified","Unified")))) %>% 
  group_by(collection,p_id) %>%
  summarise(theme_type=case_when(
    any(theme_type=="Unified") ~ "Systematisoituja", 
    any(theme_type=="Non-unified") ~ "Vain ei-systematisoituja", 
    T ~ "Ei annotointeja"), .groups="drop") %>%
  count(collection,theme_type) %>%
  collect() %>%
  mutate(collection=fct_rev(fct_relevel(str_to_upper(collection),"ERAB","SKVR","JR")),theme_type=fct_rev(fct_relevel(theme_type,"Systematisoituja","Vain ei-systematisoituja","Ei annotointeja"))) %>%
  ggplot(aes(x=collection,y=n,fill=theme_type)) + 
  geom_col() + 
  theme_hsci_discrete() + 
  xlab("Kokoelma") +
  ylab("Runoja") +
  labs(fill="Runotyyppiannotaatiot") +
  theme(legend.position="bottom") +
  guides(fill = guide_legend(reverse = TRUE)) +
  scale_y_continuous(labels=scales::number) +
  coord_flip()
```


# Spatiotemporal overview

```{r}
d <- poems %>%
  left_join(p_year %>% mutate(year=if_else(year %in% c(0L,9999L),NA,year))) %>% 
  collect() %>%
  mutate(year_ntile=ntile(year,11)) %>%
  group_by(year_ntile) %>%
  mutate(years=str_c(min(year),"-",max(year))) %>%
  ungroup() %>%
  left_join(p_loc %>% collect()) %>% 
  count(years,loc_id) %>% 
  ungroup() %>%
  left_join(locations %>% select(loc_id,name) %>% collect())
```


```{r,fig.height=11, results="asis"}
polygons %>% 
  left_join(d %>% complete(name,years)) %>%
  tm_shape() +
  tm_polygons(col='n', id='name', style='fisher', palette='plasma') +
  tm_layout(main.title="Geographical overviews by time",legend.outside.size=0.1) +
  tm_facets(by="years",ncol=4)
```

# Poem length statistics

## By collection

```{r}
poem_stats %>%
  filter(nverses<=75) %>%
  inner_join(poems) %>%
  count(collection,nverses) %>%
  ungroup() %>%
  ggplot(aes(x=nverses,y=n)) +
  geom_col(width=1) +
  facet_wrap(~collection,scales="free_y") +
  theme_hsci_discrete(base_family="Arial") +
  scale_y_continuous(labels=scales::comma_format()) +
  xlab("Number of verse lines") +
  ylab("Poems") +
  labs(title="Number of verse lines")
```

```{r}
poem_stats %>%
  inner_join(poems) %>%
  count(collection,nverses) %>%
  ungroup() %>%
  group_by(collection) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nverses<=75) %>%
  ggplot(aes(x=nverses,y=collection,fill=collection,height=prop)) +
  geom_density_ridges(stat='identity') +
  theme_hsci_discrete(base_family="Arial") +
#  scale_y_continuous(labels=scales::percent_format()) +
  xlab("Number of verse lines") +
  ylab("Poems") +
  labs(title="Number of verse lines")
```

### Poems with more than 75 verse lines

```{r}
poem_stats %>%
  inner_join(poems) %>%
  count(collection,nverses) %>%
  mutate(nl=if_else(nverses>75,n,0L)) %>%
  group_by(collection) %>%
  summarise(lines=sum(nl),proportion=sum(nl)/sum(n),.groups="drop") %>%
  arrange(desc(lines)) %>%
  gt() %>%
  tab_header(title="Poems with more than 75 verse lines") %>%
  fmt_integer(lines) %>%
  fmt_percent(proportion)
```

## By county

```{r,fig.height=11}
poem_stats %>% 
  left_join(p_loc) %>% 
  left_join(locations) %>% 
  left_join(locations,by=c("par_id"="loc_id")) %>% 
  mutate(name=if_else(type.x=="county",name.x,name.y)) %>%
  count(name,nverses) %>%
  ungroup() %>%
  group_by(name) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nverses<=40,name!="Ahvenanmaa") %>%
  collect() %>%
  mutate(name=fct_reorder(name,prop,.fun=max)) %>%
  ggplot(aes(x=nverses,y=name,height=prop)) +
  geom_density_ridges(stat='identity') +
  theme_hsci_continuous(base_family="Arial") +
#  scale_y_continuous(labels=scales::percent_format()) +
  xlab("Number of verse lines") +
  ylab("Poems") +
  guides(fill="none") +
  labs(title="Number of verse lines by county")
```

# Poem verse statistics

## By collection

### Line types

```{r}
d <- verses %>% 
  left_join(verse_poem) %>% 
  left_join(poems) %>% 
  count(collection,type) %>% 
  ungroup() %>%
  arrange(collection,desc(n)) %>%
  collect()
```


```{r}
d %>% 
  group_by(collection) %>%
  mutate(proportion=n/sum(n)) %>%
  gt() %>%
  fmt_integer(n) %>%
  fmt_percent(proportion)
```

### Verse line lengths

```{r}
d_nr_characters <- verses_cl %>%
  mutate(nr_characters=str_length(text)) %>%
  left_join(verse_poem) %>% 
  left_join(poems) %>% 
  count(collection,nr_characters) %>% 
  ungroup() %>%
  arrange(collection,desc(n)) %>%
  collect()

d_nr_words <- word_occ %>%
  group_by(v_id) %>%
  summarise(nr_words=max(pos),.groups="drop") %>%
  left_join(verse_poem) %>%
  left_join(poems) %>% 
  count(collection,nr_words) %>% 
  ungroup() %>%
  arrange(collection,desc(n)) %>%
  collect()
```

#### Verse line lengths in characters
```{r}
d_nr_characters %>% 
  filter(nr_characters<=60) %>%
  ggplot(aes(x=nr_characters,y=n)) +
  geom_col(width=1) +
  facet_wrap(~collection,scales="free_y") +
  theme_hsci_discrete(base_family="Arial") +
  scale_y_continuous(labels=scales::comma_format()) +
  xlab("Number of characters") +
  ylab("Verses") +
  labs(title="Number of characters in verse lines")
```

```{r}
d_nr_characters %>% 
  group_by(collection) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nr_characters<=60) %>%
  ggplot(aes(x=nr_characters,y=collection,fill=collection,height=prop)) +
  geom_density_ridges(stat='identity') +
  theme_hsci_discrete(base_family="Arial") +
#  scale_y_continuous(labels=scales::percent_format()) +
  xlab("Number of characters") +
  ylab("Verses") +
  labs(title="Number of characters in verse lines")
```

#### Verse lines with more than 60 characters

```{r}
d_nr_characters %>% 
  mutate(nl=if_else(nr_characters>60,n,0L)) %>%
  group_by(collection) %>%
  summarise(lines=sum(nl),proportion=sum(nl)/sum(n),.groups="drop") %>%
  arrange(desc(lines)) %>%
  gt() %>%
  tab_header(title="Verse lines with more than 60 characters") %>%
  fmt_integer(lines) %>%
  fmt_percent(proportion)
```


#### Verse line lengths in words
```{r}
d_nr_words %>% 
  filter(nr_words<=10) %>%
  ggplot(aes(x=nr_words,y=n)) +
  geom_col(width=1) +
  facet_wrap(~collection,scales="free_y") +
  scale_x_continuous(breaks=seq(0,10,by=2)) +
  scale_y_continuous(labels=scales::comma_format()) +
  theme_hsci_discrete(base_family="Arial") +
  xlab("Number of words") +
  ylab("Verses") +
  labs(title="Number of words in verse lines")
```

```{r}
d_nr_words %>% 
  filter(nr_words<=10) %>%
  uncount(n) %>%
  ggplot(aes(x=nr_words,y=collection,fill=collection)) +
  stat_binline(binwidth=1) +
  theme_hsci_discrete(base_family="Arial") +
  scale_x_continuous(breaks=seq(0,10,by=2)) +
  xlab("Number of words") +
  ylab("Verses") +
#  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Number of words in verse lines")
```

#### Verse lines with more than 10 words

```{r}
d_nr_words %>% 
  mutate(nl=if_else(nr_words>10,n,0L)) %>%
  group_by(collection) %>%
  summarise(lines=sum(nl),proportion=sum(nl)/sum(n),.groups="drop") %>%
  arrange(desc(lines)) %>%
  gt() %>%
  tab_header(title="Verse lines with more than 10 words") %>%
  fmt_integer(lines) %>%
  fmt_percent(proportion)
```

```{r}
verse_nr_words <- word_occ %>% 
  group_by(v_id) %>%
  summarise(nr_words=max(pos)) %>%
  compute_a(unique_indexes=list(c("v_id","nr_words")))

word_nr_characters <- words %>%
  mutate(nr_characters=str_length(text)) %>%
  select(w_id,nr_characters) %>%
  compute_a(unique_indexes=list(c("w_id","nr_characters")))

d <- word_occ %>%
  left_join(word_nr_characters) %>%
  left_join(verse_nr_words) %>%
  left_join(verse_poem %>% select(-pos),by=c("v_id")) %>% 
  left_join(poems) %>% 
  count(collection,nr_words,pos,nr_characters) %>%
  collect()
```

## By county

### Verse line lengths

```{r}
d_nr_characters <- verses_cl %>%
  mutate(nr_characters=str_length(text)) %>%
  left_join(verse_poem) %>% 
  left_join(p_loc) %>% 
  left_join(locations) %>% 
  left_join(locations,by=c("par_id"="loc_id")) %>% 
  mutate(name=if_else(type.x=="county",name.x,name.y)) %>%
  count(name,nr_characters) %>% 
  ungroup() %>%
  arrange(name,desc(n)) %>%
  collect()

d_nr_words <- word_occ %>%
  group_by(v_id) %>%
  summarise(nr_words=max(pos),.groups="drop") %>%
  left_join(verse_poem) %>%
  left_join(p_loc) %>% 
  left_join(locations) %>% 
  left_join(locations,by=c("par_id"="loc_id")) %>% 
  mutate(name=if_else(type.x=="county",name.x,name.y)) %>%
  count(name,nr_words) %>% 
  ungroup() %>%
  arrange(name,desc(n)) %>%
  collect()
```

#### Verse line lengths in characters

```{r}
d_nr_characters %>% 
  group_by(name) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nr_characters<=40,name!="Ahvenanmaa") %>%
  mutate(name=fct_reorder(name,prop,.fun=max)) %>%
  ggplot(aes(x=nr_characters,y=name,height=prop)) +
  geom_density_ridges(stat='identity') +
  theme_hsci_discrete(base_family="Arial") +
#  scale_y_continuous(labels=scales::percent_format()) +
  xlab("Number of characters") +
  ylab("Verses") +
  labs(title="Number of characters in verse lines")
```

#### Verse line lengths in words

```{r,fig.height=11}
d_nr_words %>% 
  filter(nr_words<8,name!="Ahvenanmaa") %>%
  mutate(name=fct_reorder(name,n,.fun=max)) %>%
  uncount(n) %>%
  ggplot(aes(x=nr_words,y=name)) +
  stat_binline(binwidth=1,scale=0.9) +
  theme_hsci_discrete(base_family="Arial") +
  scale_x_continuous(breaks=seq(0,10,by=2)) +
  xlab("Number of words") +
  ylab("Verses") +
#  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Number of words in verse lines")
```

## Number of characters in words by their position

```{r}
verse_nr_words <- word_occ %>% 
  group_by(v_id) %>%
  summarise(nr_words=max(pos)) %>%
  compute_a(unique_indexes=list(c("v_id","nr_words")))

word_nr_characters <- words %>%
  mutate(nr_characters=str_length(text)) %>%
  select(w_id,nr_characters) %>%
  compute_a(unique_indexes=list(c("w_id","nr_characters")))

d <- word_occ %>%
  left_join(word_nr_characters) %>%
  left_join(verse_nr_words) %>%
  left_join(verse_poem %>% select(-pos),by=c("v_id")) %>% 
  left_join(poems) %>% 
  count(collection,nr_words,pos,nr_characters) %>%
  collect()
```

```{r}
d %>%
  group_by(collection,nr_words,pos) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nr_words>=2L,nr_words<=5L) %>%
  mutate(nr_words=as_factor(nr_words),pos=as_factor(pos)) %>%
  uncount(n) %>%
  ggplot(aes(x=nr_characters,y=nr_words,fill=pos)) +
  stat_binline(binwidth=1) +
  facet_grid(collection~pos,labeller = labeller(pos=label_both)) + 
  xlab("Number of characters in word") +
  ylab("Number of words in verse") +
  labs(
    title="Number of characters in words by their position",
    subtitle="According to length of verse and collection"
    ) +
  guides(fill="none") +
  theme_hsci_discrete(base_family="Arial")
```

```{r}
d %>%
  group_by(collection,nr_words,pos) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup() %>%
  filter(nr_words>=2L,nr_words<=5L) %>%
  mutate(nr_words=as_factor(nr_words),pos=as_factor(pos)) %>%
  uncount(n) %>%
  ggplot(aes(x=nr_characters,y=pos,fill=nr_words)) +
  stat_binline(binwidth=1) +
  facet_grid(collection~nr_words,labeller = labeller(nr_words=label_both)) + 
  xlab("Number of characters in word") +
  ylab("Position") +
  labs(
    title="Number of characters in words by their position",
    subtitle="According to length of verse and collection"
    ) +
  guides(fill="none") +
  theme_hsci_discrete(base_family="Arial")
```
